### 2013-06-10 paper/jneurosci2013.tex
Second version of the paper, with figures.

\documentclass[12pt]{article}
% Change "article" to "report" to get rid of page number on title page
\usepackage{amsmath,amsfonts,amsthm,amssymb}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{extramarks}
\usepackage{chngpage}
\usepackage{soul}
\usepackage[usenames,dvipsnames]{color}
\usepackage{graphicx,float,wrapfig}
\usepackage{listings}
\usepackage{framed}
\usepackage{inconsolata}
\usepackage{cite}

% Stuff to change, you know, if you want.
\setlength{\parindent}{1cm}
\setlength{\parskip}{3pt}

% In case you need to adjust margins:
\topmargin=-0.45in      %
\evensidemargin=0in     %
\oddsidemargin=0in      %
\textwidth=6.5in        %
\textheight=9.0in       %
\headsep=0.25in         %

\pagestyle{empty}

\newcommand{\scalefigone}[3]{
  \begin{figure}[ht!]
    % Requires \usepackage{graphicx}
    \centering
    \includegraphics[width=#2\columnwidth]{#1}
    \caption{#3}
    \label{#1}
  \end{figure}}

\setlength\fboxrule{1pt}

\graphicspath{{../figures/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Make title
\title{}
\date{}
\author{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{center} \begin{Large}
  \textbf{A task-state model of medial PFC \\ can perform top-down
    control}
\end{Large} \end{center}

\begin{center}
  Trevor Bekolay, Benjamine Liu, Nandakumar Narayanan, Chris
  Eliasmith, Mark Laubach
\end{center}

\begin{abstract}
  Many hypotheses on the function of medial prefrontal cortex (mPFC)
  have been put forth,
  including action-outcome learning, ???More.
  We adopt a more general approach and posit that
  the mPFC tracks task state,
  which can be usd to accomplish previously hypothesized functions.
  We use as a model rodents performing a simple reaction time (RT) task.
  First, we show that mPFC activity
  tracks task state in the simple RT task,
  and that neural integration is a core mechanism
  for accomplishing task state tracking.
  Second, we show that a working memory model
  using neural integrators closely matches
  rodent neural activity.
  Finally, we propose a neural circuit model
  that imposes top-down control on a control system
  as a proof-of-concept that the task-state model
  can explain the myriad of functions ascribed
  to the medial PFC.
\end{abstract}

\section{Introduction}

mPFC function background

mPFC models background

We propose that the function of the mPFC is to
track task states.
These task states can be very high-dimensional,
if, for example, visual features are necessary
for performing a particular task.
The other functions that have been ascribed
to the mPFC are accomplished
by other areas, but use the task state
information in the mPFC to perform their function.

As an application of this theory,
we investigate a simple reaction-time task.
The remainder of the paper is organized as follows.
First, we review experimental data from rodents
doing the reaction-time task.
The task state space being tracked is determined
by a dimensionality reduction technique
(principal component analysis).
Second, we review a spiking neural model
of working memory and
analyze it during the simple reaction-time task.
Third, we analyze the two-dimensional dynamical system
that makes up the task state space and
show how it changes as a result of performing the task.
Fourth, we create a neural control system
that can perform the simple reaction-time task
and show how the mPFC model can exert top-down control
on that control system.
Finally, we give some examples of how the model
can be manipulated in order to predict
neural and behavioural effects in certain pathologies.

\section{Experimental data}

Narayanan \& Laubach (2009) recorded the medial PFC
of rodents performing a simple reaction-time task
(figure ???).
In this task, the subject presses and holds down the lever
for a delay (the ``foreperiod'';
1 second for the data in this study).
After the foreperiod, an auditory tone (stimulus) is played.
The subject then has 0.6 seconds (the response window)
in which to release the lever.
If the lever is released in that time,
the subject is rewarded.
If the lever is released during the foreperiod,
it is called a premature error,
and is penalized by a time-out.
If the lever is not released within the response window,
it is called a late error,
and is also penalized by a time-out.

Narayanan \& Laubach used principal component analysis
in order to determine the important signals
being represented in the mPFC.
As shown in figure ???,
the first two principal components were significant,
and appear to track
1) if the subject was performing the task, and
2) the relative amount of time the subject has been
performing the task.
Narayanan \& Laubach also found that the neural activity
in the mPFC changed significantly depending on
the outcome of the previous trial;
the first two principal components
change significantly after errors,
and appear to contain a trace that an error occurred.

Interestingly, in the post-correct case,
the first principal components
is very high correlated with the integral
of the second principal components,
and similarly the second principal component
with the integral of the first.
In the post-error case,
these correlations are significantly weaker.
This result suggests that
the mPFC is performing some kind of integration.

\section{Coupled neural integrators}

One model of neural integration is
Singh \& Eliasmith (2006).
Their model was created to explain working memory activity
in PFC during a vibrotactile discrimination task.
It employs two-dimensional neurons in integrative populations
in order to show time-varying activity during the task.
Due to the identification of two strong components
in the rodent data,
we hypothesized that the Singh \& Eliasmith model
could be used to model the simple RT task as well.

There are two different realizations
of this model, as shown in figure ???.
In one, a single population of adaptive
leaky integrate-and-fire neurons tracks both
integrative dimensions, as well as two control dimensions.
In the other, two coupled populations
of adaptive leaky integrate-and-fire neurons
each track one integrative dimension, as well
as one control dimension.
The integrative dimensions are projected
to another population that does not explicitly
represent the control dimensions.
In both realizations of the model,
input to the integrators is provided
by motor actions, and are controlled
by the task outcome.
The idea here is that a motor action
(in this case a lever press)
starts the presistent activations
of one of the integrative dimensions.
That dimension provides input to the other integrative dimension.
Task outcomes (in this case either reward
or time-out penalties) affect the control dimensions,
either stopping all integration
in the reward case or
reversing the integration in the penalty case.

These two different model configurations
make an important distinction between
what is being represented in the brain
area being analyzed.
In our case, should the single population model
match the neural data,
it predicts that the task information
as well as the action and outcome information
is being represented in the mPFC.
If the coupled population model matches
the neural data,
it predicts that only the task information
is being represented, and action and outcome
information only affects the mPFC's
task state.

As can be seen in figure ???,
the coupled integrator model matches
the neural data closely, while the single population model
does not.
In the single population model,
there is highly modulated activity
around the time of reward delivery;
this modulation is not seen in the data.
Therefore, it can be concluded that,
at least for the rodent in a simple RT task,
outcome information is not explicitly
represented in the mPFC,
though the outcome does affect the task
information reflected in mPFC activity.
We will analyze the coupled integrator
realization of the Singh \& Eliasmith
model in the subsequent sections.

\section{Task state space analysis}

If the role of the mPFC is to track the task-state,
then we can view mPFC activity as a trajectory
in the task state space.
This is done in figure ??? for the experimental data,
and the simulated data.

In the post-correct case,
the trajectory through task space is cyclical.
In other words,
when the subject is performing the task properly,
the mPFC travels in a predictable loop.
In the post-error case,
the trajectory through task space is less predictable.

The dynamics of this task state space
can be approximated based on these trajectories,
and on how the Singh \& Eliasmith model was created.
These dynamics are captured in the following equation.

\begin{equation*}
  \begin{bmatrix}
    \dot{x_1} \\ \dot{x_2}
  \end{bmatrix} = \begin{bmatrix}
    -\beta R & 0 \\ \alpha & -\beta R
  \end{bmatrix} \begin{bmatrix}
    x_1 \\ x_2
  \end{bmatrix} + \begin{bmatrix}
    1 & -1 \\ 0 & 0
  \end{bmatrix} \begin{bmatrix}
    A \\ P
  \end{bmatrix} 
\end{equation*}
where $A$ is the action (lever press),
$R$ is the reward, and $P$ is the penalty.
Figure ??? shows the result of directly simulating
the trajectory of this dynamical system.

The Singh \& Eliasmith model uses the methods
of the Neural Engineering Framework (NEF),
which can simulate arbitrary dynamical systems.
The dynamical system defines how the model
encodes information in the populations of neurons,
and how those encoded values are transformed
through connections between populations.
Because we know exactly how the information
is encoded and transformed, we can also
decode that information directly;
this results in figure ???,
which looks like a noisy version
of the mathematical model.
It is important to note that this decoding process
was done on the same neural data that
results in the principal components in figure ???;
therefore, we can say with some certainty
that the mathematical model is closely emulating
the experimental mPFC.

\section{Top-down control of motor actions}

Previous work by Laubach and colleagues (and other groups)
has implicated mPFC as a top-down controller
on the motor cortex in delay tasks like simple RT.
Their work suggests that the mPFC directly inhibits
the motor cortex.
In this paper, we propose instead that the mPFC
can either inhibit, directly excite,
or modulate other activity being projected
to the motor cortex depending on
the part of task-state space the mPFC is in.

In order to demonstrate that this is possible,
we created a simple stimulus-response control system
using the methods used in Singh \& Eliasmith, 2006
(i.e., the NEF; see figure ???).
Briefly, the system is composed of three populations
of adaptive leaky integrate-and-fire neurons.
One population presses and releases the lever,
one population tracks if the lever is currently being held down,
and one population is sensitive to the auditory stimulus,
and stimulates release neurons if the lever
is currently being held down.
This system is trivial to build,
and can perform the simple RT task
with little difficulty (see figure ???).

Top-down control can be accomplished
because of the predictability with which
the mPFC traverses the task state space.
If the subject continues to perform the task correctly,
the release should happen at the same part
of the state space (see figure ???);
if the subject forces a release at the appropriate
part of the state space, it can, in fact,
release the lever faster than it would be able to
if it waits for the stimulus.
In fact, if the foreperiod is completely
predictable, then even devoid of a stimulus,
the subject could perform this task
solely through a forced release mediated
by the mPFC.

Through a similar argument, the mPFC
can also cause the subject to be more
cue-sensitive after errors occur.
When an error occurs, the mPFC is driven to a
low part of the state space.
When the initial press occurs, it may not
be driven to the top part of the state space
as it should; if that occurs (i.e.,
if the subject is performing this task
in the wrong part of the state space),
the subject can attempt to become more cue-sensitive
(see figure ???).

This top-down control can be exerted
on the stimulus-response motor control system
with the network configuration depicted in
figure ???. Briefly, top-down control
is mediated by directly stimulating
the population that is sensitive to the auditory stimulus.
When the mPFC is in the force-release zone,
then these neurons will be excited;
this can either make the population react quicker to
the stimulus, or can directly cause a release
even in the absence of a stimulus.
When the mPFC is in the cue-sensitive zone,
these neurons are stimulated in such a way that
they react quicker and stronger to the stimulus,
but cannot directly cause a release
in the absence of a stimulus.

\section{Task-state space dynamics and errors}

Looking at the dynamics of the task state space
at each part of a simple RT task trial
(figure ???)
gives a full picture of what the task-state
theory of mPFC function predicts is occurring
during the simple RT task.

At the start of a trial,
the system is driven upwards towards [1.0, 0.0].
This sets the initial state for the task.
The subject may have been doing another task
before initiating this RT task;
by resetting to the same initial state,
the subject's behaviour can proceed predictably.
As will be seen, however, after errors,
the system may be sufficiently far from this
initial state that it cannot quite get to the initial state,
and instead operates in an unpredictable part of the task space.

In a correct trial (b), the passive dynamics of the system
push the state to the right
at a predictable (i.e., learned) rate.
The release happens when the state of the system
is sufficiently far to the right (i.e., $x_1$ is high).
If this release happens properly, reward is delivered,
which pushes the system back to the origin.
Then, in the absence of any task-related dynamics,
the system will stay approximately at rest
because the passive dynamics of the system are very weak
at the origin point.

A premature error can be caused by some other factor
(i.e., a motor system mistake) or by the passive dynamics
of the system being too strong in the part of the state space
that the system is driven to once the subject presses the lever.
In the case of some other factor,
the system will receive the negative outcome feedback
before reaching the right side of the state-space,
and be driven to the bottom.
In the case of fast integration,
the system will reach the right side of the state-space,
force a release, and then receive the negative outcome feedback
and be driven to the bottom.
The exact state after receiving the negative feedback
can be used to figure out the cause of the error,
though after some time in the lower part
of the state space, the system will be
driven to the left,
and all that can be decoded
is the notion that an error occurred.

A late error can be caused by either another factor
not responding to the top-down control properly or in time,
by right-ward drift happening too slowly,
or by the system not being driven
to the appropriate place upon pressing down the lever
and drifting slowly as a result.

\section{Behavioural results}

While there are a number of tunable parameters,
(e.g., $\alpha$ and $\beta$ from equation ???),
there are many more parameters that are randomly selected
from probability distributions.
Neuron parameters (neuronal gain, encoding vectors,
adaptation constants, etc.) are all randomized,
and the results of those randomizations are used
to solve for appropriate connection weights.
As such, there is a large amount of variation
between instances of this model.
For that reason,
we generate several instances of the model
and run the model for over forty trials
to get an idea of how
that particular instance of the model behaves.

Figure ??? summarizes the behavioural results
from the experimental rodents, the stimulus-response model
alone, and the stimulus-response model with top-down
control by a coupled integrator model.
The experimental rodents ???.
The stimulus-response model by itself has a relatively slow
reaction time (median ???, $\pm$ ???),
but very high accuracy (??? correct, ??? late, 0\% early).
With top-down control, some models were able to achieve
extremely fast reaction times (e.g., ???).
On average, however, the median reaction time was
??? $\pm$ ???. Note that the variance in reactions times
was much higher with top-down control,
and the accuracy was in general lower (??? correct, ??? late,
??? early).

\section{Model manipulations}

As a spiking neuron model,
this theory of mPFC function is uniquely able to be
manipulated; manipulations at the neural level
can be examined at any point, from the single-cell
to the behavioural level.

One possible manipulation is to
ablate a portion of recurrent synapses,
which could model the effect of brain trauma or aging.
The effects on the mPFC's ability to track the
task state space can be seen in figure ???.
%%% MORE

Since we also know the mathematics behind the mPFC activity,
we can instead manipulate the mathematical model
and make detailed predictions about neural activity.
For example, it is hypothesized that in aging,
recurrent connections become weaker, and therefore
the persistent activity that is seen in working memory models
cannot be maintained for long periods of time.
This can be modeled by including a small damping term, $d < 0$,
in equation ???.

\begin{equation*}
  \begin{bmatrix}
    \dot{x_1} \\ \dot{x_2}
  \end{bmatrix} = \begin{bmatrix}
    -\beta R + d & 0 \\ \alpha & -\beta R + d
  \end{bmatrix} \begin{bmatrix}
    x_1 \\ x_2
  \end{bmatrix} + \begin{bmatrix}
    1 & -1 \\ 0 & 0
  \end{bmatrix} \begin{bmatrix}
    A \\ P
  \end{bmatrix} 
\end{equation*}

In addition to seeing the lower-level effects
of these two manipulations,
we can also make the same manipulations in the full
top-down model. As can be seen in figure ???, % STUFF

\subsection{Predictions}

??? collect together predictions

\section{Discussion}

??? Talk about the results

On interesting feature of the model is its stochastic nature.
While it can be though to represent individual differences,
it is also a useful notion to consider that the model
represents one snapshot during a subject's
learning of the task.
There is no learning in this model;
while the connection weights are solved for ``optimally,''
due to the large amount of noise in the system,
learning would be very likely to fine-tune
the mPFC's ability to traverse and
be sensitive to parts of the task-state space.
Therefore, the poor performance of, for example,
top-down model instance ??? can be thought of as
a subject whose recurrent connection is too strong
and therefore is a poor performer in the task,
or it could be thought of as a subject
who has not yet learned how to properly time
the interval, and after more trials would,
with learning, be able to make less premature errors.

Stepping back, the overarching theory of the paper,
that the mPFC tracks the task state space, ??? more

The theory sets up the possibility for many arbitrary tasks.
Basically, each task has $N$ important dimensions.
The passive dynamics are likely a combination
of integrative and non-integrative dimensions,
enabling for flows in many directions of the space.
The mPFC is drive to specific parts of the state space
based on the sensory environment (e.g.,
being pushed to the top of the state space when lever pressing).
The mPFC then learns to be sensitive to salient
parts of the state space (e.g.,
the force release zone).

\clearpage

\appendix

\section{Materials and methods}

\subsection{Simple reaction time task}

In both the empirical and simulated data,
subjects and models performed the same simple reaction time task.
At the start of a trial,
subjects press down a lever.
A timer tracks the time since the lever was pressed down,
and after an experimenter-specified time period,
called the \textit{delay period} or \textit{foreperiod},
an auditory cue signals that the subject must respond
by releasing the lever.
If the lever is released within 600 milliseconds,
the trial is considered a success
and reward is delivered.
If the lever is released before the auditory cue
or after the 600 millisecond response window,
then the trials is considered a failure
and subjects are given a time out in which
lights are extinguished.
Lever presses during this timeout period
lengthen the timeout period.

Empirically, this study is done with rats
in a MedPC box containing a lever,
house lights, and a spout for delivering water reward.
Rats were trained by first
associating reward with lever presses,
and then introducing progressively longer
delay periods until they reached the goal delay
of 1 second.
They were considered to have learned the task
when they performed the task correctly $>60\%$
of the time.

Theoretically, this study was performed
in the Nengo simulation environment.
Subjects were networks of leaky integrate-and-fire
neurons, as described below.
The environment was modeled using a
\texttt{simplenode}, which allow arbitrary code
to interact with a neural model.
In this case, the \texttt{simplenode} implemented
a finite state machine to track task state,
and scalar values to indicate the state of
house lights, lever position, auditory cue, and reward.

\subsection{Empirical summary}

The empirical study (described in more depth in ?) %???cite???)
was done with twelve male Long-Evans rats (3-4 months old; Harlan).
Rats were water restricted, with food always available,
and consumed 10-15 ml of water during each
behavioural session.

Arrays of recording electrodes were implanted
in dmPFC ($n = 8$), motor cortex ($n = 3$), or both ($n = 1$).
Electrodes were 50 $\mu$m stainless steel wires
with 250 $\mu$m between wires.
Arrays were in configurations of
4 $\times$ 4 or 3 $\times$ 3 $\times$ 2.
Microwire arrays were targeted at dmPFC
based on the following coordinates from bregma:
anteroposterior [AP]: $+$3.2mm;
mediolateral [ML]: $\pm$1.4mm;
dorsoventral [DV]: $-$3.6mm at 10$^\circ$ in the frontal plane.
Motor cortex was targeted with
AP: $-$0.5mm; ML: $\pm$2.5-3.5mm;
DV: $-$1.5mm at $-25^\circ$ in the frontal plane.
Histology was performed to confirm electrode locations
(???see fig???).

Ensemble recordings were done using a
Many Neuron Acquisition Program (Plexon, Dallas, TX).
Spike sorting using principal component analysis (PCA)
and waveform shape identified 414 single units
whose spikes were included in this study.

\subsection{Theoretical}

\subsubsection{Double-integrator model}

The Singh \& Eliasmith model (2006).
This model is created using the methods of the
Neural Engineering Framework.
The model uses neural integration in order to explain
neural data in a working memory task.
The paper proposes two candidate models
that can create the desired functional effects
with different neural dynamics.

In the single population model, a single population of
adaptive leaky integrate-and-fire neurons represents
a four-dimensional signal.
A recurrent connection causes the signals
being represented in two of the dimensions
to integrate; the connection
also causes one of the integrative dimensions
to drive the second integrative dimension.
The remaining two dimensions are used
to control the integrative dimensions,
returning them to some baseline activity if necessary.

In the coupled population model, two populations of
adaptive leaky integrate-and-fire neurons represent
two two-dimensional signals.
Recurrent connections on each population cause
one of the represented dimensions to be integrated,
while the other dimension controls the integration.
An explicit connection between one integrative population
drives the integrator in the other population.
Both of the integrative dimensions are then
projected to a final two-dimensional population,
which only represents the integrative signals.

In this paper, we hypothesize that the mPFC
is a realization of one of these two models.
The integrators are driven by the actions taken by the subjects
(e.g., lever pressing),
and are controlled by the outcomes
(e.g., reward delivery).

\subsection{Analysis}

Both empirical and theoretical experiments
produce data in the form of spike trains
and event timestamps.
The files produced by NeuroExplorer and Nengo
were read by a customized version of Neo,
a Python package for reading neurophysiology data files.
The subsequent analysis was done
with custom written scripts in Python
using the NumPy and SciPy libraries for numerical analysis
and Matplotlib for plotting.
All analysis scripts are freely available
at ?. % github URL

\subsubsection{Behaviour}

In both the empirical and theoretical studies,
the following relevant event times are recorded:
lever press, lever release, reward delivery,
trigger stimulus onset, and house light extinguishing.
Sequences of event times can be used to
identify the result of each trial,
as shown in Figure ?.
Filtering the set of event timestamps
for these sequences identifies each trial outcome.

Reaction time distributions are also analyzed
for correct trials.
Reaction time is defined as the difference between
trigger stimulus onset and lever release.

\subsubsection{Neural}

Principal component analysis (PCA)
is used to identify the important features
in spike trains around certain events.
The previously mentioned techniques
were used to isolate an event of interest.
For each neuron, spike trains from 1 second prior to
2 seconds after each instance of an event are isolated.
Each perievent spike train is binned in 1 ms bins and is then
convolved with a Gaussian filter ($\sigma = $ 25 ms;
PCs are consistent with many other $\sigma$ values).
The spike trains were slightly extended
in order to eliminated edge artifacts.
These spike density functions
are normalized to z-scores,
and then averaged over all trials to
produce a matrix in which each row
the normalized average response of a neuron
over the perievent epoch.
PCA was performed on that matrix
using singular value decomposition.
The resulting principal components
were also normalized to z-scores.
The amount of variance accounted for by each principal component
is computed as the square of the component's eigenvalue
over the sum of all squared eigenvalues
(i.e., $s_i^2 / \sum s^2$).

In all analyses, neurons not significantly modulated
during the perievent epoch were omitted.
Specifically, we omitted neurons with average firing rates
under 1 Hz and whose firing rates were non-stationary
across trials
(determined by a Wald-Wolfowitz runs test with $|Z| > 4$).

\clearpage

\section{Figures}

\scalefigone{f1.png}{1.0}{~}

A summary of previous experimental findings. Data from JNP 2009.

A: The simple reaction time task used in the experiment.
The subject presses and holds down the lever
for a delay (also called the foreperiod);
1 second is used in this study.
After the foreperiod, an auditory tone (the stimulus) is played.
The subject then has 0.6 seconds (called the response window)
in which to release the lever.
If the lever is released in that time,
the subject is rewarded with juice.
If the lever is released during the foreperiod,
it is called a premature error,
and is penalized by a 5 second time-out. % NOTE: HOW LONG IS TIMEOUT?
If the lever is not released within the response window,
it is called a late error,
and is also penalized by a 5 second time-out.

B: When lever release events occurred.
Each trial is aligned such that the press event
occurs at time $t=0$.
All of the trials shown are correct trials.
Black dots represent correct trials
that were preceded by correct trials.
Red dots represent correct trials
that were preceded by error trails
(either premature or late).

C: Peri-event spike rasters and histograms
showing different neural activity
based on the previous outcome.
In all six plots, black represents activity
when the last trial was correct, and red
represents activity when the last trial was an error.
The shaded grey areas note where
the neural activity is significantly different
between the two conditions (post-correct and post-error).
In the top row, there is elevated neural activity
before the press is initiated, following an error.
In the middle row, there is elevated neural activity
before the press is initiated (both)
and during the foreperiod (right)
following correct trials.
In the bottom row, there is elevated neural activity
during the whole trial (left)
and after the trial (right)
following error trials.

D: Linear regression with a GLM was performed
on the neural activity.
Before the press is initiated,
18.7\% of neurons encoded the previous outcome,
compared to less than 10\% that
encoded the current reaction time. % WHAT DOES INTERACTION MEAN HERE?

E: Z-statistics showing which neurons were significantly
modulated by the previous outcome, and when.
(Top) Sorted by the absolute value of the Z-statistic.
The bottom 18.7\% of the rows have high average Z-scores
($p > 0.05$?). % WHAT SAYS THOSE ARE SIGNIFICANT?
(Bottom) Sorted by the peak of the Z-statistic.
The line is close to linear, suggesting that the
previous outcome information is being encoded over time
by the entire population. % DOES IT SUGGEST THAT?

\clearpage

\scalefigone{f2.png}{1.0}{~}

A summary of the principal component analysis done
in JNP 2009;
the motivation for investigating Singh \& Eliasmith 2006.

A: A graphical summary of principal component analysis.
Nomarlized peri-event neural data,
in the form of the z-scores of instantaneous firing rates,
is organized in a matrix, with each row
being the z-scored firing rates
of a single neuron on a single trial.
Singular value decomposition is performed on the matrix,
resulting in a matrix such that
the number of columns (time bins) is the same.
The rows are now ordered such that the first row
contains largest eigenvector,
which represents the value when the original
data is projected onto the axes of highest variance.

B: The fraction of variance explained
by each singular value
when PCA is performed on the post-correct neural data
(top), and the post-error data (bottom).
In both cases, the two eigenvectors with the highest
singular values account for nearly 50\% of variance.
The next two eigenvectors also account for an amount
of variance higher than would be expected when
we interpolate from the smaller
singular values.

C: A summary of the loadings of the two top
eigenvector on each neuron.
If a neuron were encoding both the first and second
eigenvectors to the same degree,
we would expect a horizontal line.
The varied lines criss-crossing each other
suggests that each individual neuron
has different sensitivities
to the first and second principal components.

D: The top three principal components
for the post-correct trials (black)
and the post-error trials (red).

In the post-correct case,
the first principal component is generally high
during the foreperiod of the task, with bumps
before and after the foreperiod, and is low elsewhere.
The second principal component appear to ramp up
before the task, and down during the task.
The third principal component is similar
to the first.

In the post-error case,
the first principal component is generally high
before the press event, mostly constant during
the foreperiod, and low after the trial.
The second principal component is generally high
during the foreperiod of the task,
and closely resembles the first component
in the post-correct case.
The third principal component ramps down
during the task state, and closely resembles
the second component in the post-correct case.

E: (Top) Plotting the cumulative sum (integral)
of the first principal component closely matches
the second principal component, and vice-versa
($R^2 = 0.904$ and $R^2 = 0.939$ respectively),
in the post-correct case.
(Bottom) In the post-error case,
the first two principal components are no longer
cumulative sums of each other
($R^2 = 0.639$ and $R^2 = 0.676$).
This points to neural integration
as a potential mechanism explaining mPFC activity.

F: Normalized spike-density functions
for all of the neurons analyzed in the post-correct case,
organized by the loading
(left) on the first principal component, and
(right) on the second principal component.
These clearly show neurons whose activity
matches the calculated principal components.

\clearpage

\scalefigone{f3_blocks.pdf}{1.0}{~}

The Singh \& Eliasmith model (2006).
This model is created using the methods of the
Neural Engineering Framework.
The model uses neural integration in order to explain
neural data in a working memory task.
The paper proposes two candidate models
that can create the desired functional effects
with different neural dynamics.

(Left) A single-population model.
In this model, a single population of
adaptive leaky integrate-and-fire neurons represents
a four-dimensional signal.
A recurrent connection causes the signals
being represented in two of the dimensions
to integrate; the connection
also causes one of the integrative dimensions
to drive the second integrative dimension.
The remaining two dimensions are used
to control the integrative dimensions,
returning them to some baseline activity if necessary.

(Right) A coupled-population model.
In this model, two populations of
adaptive leaky integrate-and-fire neurons represent
two two-dimensional signals.
Recurrent connections on each population cause
one of the represented dimensions to be integrated,
while the other dimension controls the integration.
An explicit connection between one integrative population
drives the integrator in the other population.
Both of the integrative dimensions are then
projected to a final two-dimensional population,
which only represents the integrative signals.

In this paper, we hypothesize that the mPFC
is a realization of one of these two models.
The integrators are driven by the actions taken by the subjects
(e.g., lever pressing),
and are controlled by the outcomes
(e.g., reward delivery).

\clearpage

\scalefigone{f3_pca.pdf}{1.0}{~}

A comparison between the two model realizations
from Singh \& Eliasmith in this task.

A: Raw spike rasters generated by the two models.
Modulation around the foreperiod can clearly be seen.
Differences between the post-correct and post-error
conditions can also be determined,
though the difference is more visually discernible
in the coupled model.

B: Principal component analysis
on the simulated data (right) shows similar trends
as the experimental data (left).
% MORE DETAIL

C: Normalized spike-density functions,
organized by loadings on PCs (as in the experimental data)
show neurons that clearly follow the PCs.
Neurons that do not follow the PCs
can also be visually identified.

\clearpage

\scalefigone{f4_traj.pdf}{1.0}{~}

Plotting the state space (phase space) of the data
allows us to reason about what the principal components
may be representing in general.

A: The experimental data's trajectory
through the phase space formed by the first two
principal components.

B: The simulated data's trajectory
through the phase space formed by the first two
principal components.

C: A mathematical model of the trajectory
through state space.
This model is defined by a two-dimensional
dynamical system that is affected
by the actions taken by some outside system,
and the outcomes of those actions.

D: When the mathematical model is implemented
in a spiking neural model using the NEF,
we can look at the decoded activity in the
two integrative dimensions,
instead of relying on the PCs

\clearpage

\scalefigone{f5_tdcontrol.pdf}{1.0}{~}

Previous work by Laubach et al. (and other groups)
has implicated mPFC as a top-down controller
on the motor cortex in delay tasks like simple RT.
Their work suggests that the mPFC directly inhibits
the motor cortex; this paper's hypothesis
is more general, and proposes that the mPFC
can either inhibit, directly excite,
or modulate other activity being projected
to the motor cortex
depending on the state of the mPFC.

A: A schematic of how top-down control
could be exerted on a simple stimulus-response
motor control system.
The mPFC state is tracked in this two-dimensional system.
In the green area, which is entered
by the black trajectory
(exemplifying correct trials),
a release is forced to occur.
In the red area, which is entered
by the red trajectory
(exemplifying error trials),
the subject is not performing the task well,
and enters a cue-sensitive state,
in which gain is applied
to the auditory cue stimulus input.

B: The simple stimulus-response motor control system
on which top-down control is being exerted.
This system is shown to perform the simple RT task
with high accuracy,
but with a relatively slow reaction time
(see Fig ???).
This is due to the system waiting until the cue
to start the process of initiating the release.

Briefly, the stimulus response system is made up
of three populations of adaptive leaky integrate-and-fire neurons
that interact with the environment
(explicitly modeled in the Nengo simulation environment).
At the start of a trial, the press-sensitive neurons
in the \textit{Press/Release} population become active.
This drives the \textit{Holding} integrator,
which will maintain a relatively high amount of activity.
This activity is projected to the \textit{Trigger} neurons,
which are then primed to cause a release.
When the trigger stimulus is delivered by the environment,
it further activates the \textit{Trigger} neurons,
which require both inputs in order to activate
the release-sensitive neurons
in the \textit{Press/Release} population,
causing the release of the lever
modeled by the environment.

C: The combined stimulus-response motor control system,
with the Singh \& Eliasmith coupled-integrators
providing top-down control.
Top-down control is exerted through the \textit{Trigger}
population, allowing it to explicitly force a release
by providing input in both dimensions,
or sensitizing and desensitizing the population
to the trigger stimulus
by providing input only in the same dimension
that the \textit{Holding} population projects to.
Details on the coupled-integrator model
can be found in figure ???.

\clearpage

\scalefigone{f6_dynamics.pdf}{1.0}{~}

Breaking down the mathematical model
into what happens in the three cases gives
an intuitive explanation for how this dynamical system operates.

A: The start of all trials.
In all cases, no matter what the initial state,
the system is drive upwards,
close to [1.0, 0.0]. This sets the initial state
for the task.

B: In a correct trial, the passive dynamics of the system
push the state to the right
at a predictable (i.e., learned) rate.
The release happens when the state of the system
is sufficiently far to the right (i.e., $x$ is high).
If this release happens properly, reward is delivered,
which pushes the system back to the origin.
Then, in the absence of any task-related dynamics,
the system will stay approximately at rest
because the passive dynamics of the system are very small
at the origin point.

C: A premature error can be caused by some other factor
(i.e., a motor system mistake) or by the passive dynamics
of the system being too strong in the part of the state space
that the system is driven to once the subject presses the lever.
In the case of some other factor,
the system will receive the negative outcome feedback
before reaching the right side of the state-space,
and be driven to the bottom.
In the case of fast integration,
the system will reach the right side of the state-space,
force a release, and then receive the negative outcome feedback
and be driven to the bottom.
The exact state after receiving the negative feedback
can be used to figure out the cause of the error,
though after some time in the lower part
of the state space, the system will be
driven to the left,
and all that can be decoded
is the notion that an error occurred.

D: A late error can be caused by either another factor
not responding to the top-down control properly or in time,
by right-ward drift happening too slowly,
or by the system not being driven
to the appropriate place upon pressing down the lever
and drifting slowly as a result.

\clearpage

\scalefigone{f7_behaviour.pdf}{1.0}{~}

Summary of behaviour in experimental subjects,
simulated subjects with only the stimulus-response system,
and simulated subjects with the stimulus-response system
under top-down control.

A: Experimental results.
% GIVE TRENDS AND STUFF

B: For the stimulus-response system,
% GIVE TRENDS AND STUFF

C: For the system with top-down control,
% GIVE TRENDS AND STUFF

\clearpage

\scalefigone{f8_manipulations.pdf}{1.0}{~}

The model can be manipulated in several ways
in order to investigate the effects
of various pathologies.

A: By including a damping term in the dynamical system,
we impair the system's ability to maintain its state.
This can be visualized as an attractive force
pulling the system towards the origin.
This makes it more difficult or impossible
to confidently exert top-down control,
depending on the amount of damping.
This effect can be modeled by degrading
the recurrent connection weights by a small amount,
causing the integrative populations to drift
back to the origin.
The amount of weight degradation affects
how quickly the population drifts back to the origin.
This model realization of this damping
closely resembles what happens in
normal aging: recurrent NDMA connections
lose synaptic strength.
The model predicts that the new trajectory
in the state space would more closely resemble
those in the damped mathematical model.

B: The simulated model can
instead be manipulated by randomly deleting
recurrent synaptic connections.
When this is done,
the model takes the trajectories shown.
This predicts what may occur in the mPFC
in the case of targeted lesions,
or general trauma such as a concussion.

C: The behavioural effects of these two
model manipulations when they are done
on the full model with top-down control.
% GIVE SOME EXPLANATIONS

\end{document}