### 2013-06-10 paper/coverletter.tex
Fourth version of the paper, with figures.
- This is the version submitted to JNeuroscience

% LaTeX resume using res.cls
\documentclass{letter}
\usepackage[T1]{fontenc}
\usepackage{charter}
\usepackage{inconsolata}
\usepackage{graphicx}

\begin{document}

Dear Editor,

Please accept for consideration our paper entitled
``A spiking neural integrator model of the adaptive
control of action by the anterior cingulate cortex.''

This paper presents a mathematical hypothesis
to explain the role of the anterior cingulate cortex (ACC)
in the control of action.
This hypothesis is tested by implementing
it in a spiking neural network model,
and it is validated by showing it qualitatively
and quantitatively matches both neural and
behavioral data recorded from rodents.
We believe that the proposed model is the most
detailed account of ACC function
to date. We suggest that our framework can
be extended to explain a variety of
ACC functions we have not directly explored.

This study is the result of a close collaboration
between the Centre for Theoretical Neuroscience
at the University of Waterloo
and the John B. Pierce laboratory in New Haven.
We believe that this combination of
experimental and theoretical expertise
is one of the strengths of this paper.

Sincerely, \\
Trevor Bekolay, Mark Laubach, Chris Eliasmith

\end{document}