### 2013-11-14 paper/rebuttal.tex
Revised version submitted Nov 29
- Responded to JNeurosci reviews
- Now using doit to analyze, plot etc
- Many other improvements; see the diff

\documentclass[11pt,paper=letter]{scrartcl}
\usepackage[english]{babel}
\usepackage{amssymb,amsmath}
\usepackage{namedplus}
\usepackage[font=sffamily]{quoting}
\usepackage{helvet}
\usepackage[backgroundcolor=yellow,linecolor=yellow]{todonotes}
\usepackage{hyperref}
\hypersetup{
  colorlinks = true, %Colours links instead of ugly boxes
  urlcolor = blue, %Colour for external hyperlinks
  linkcolor = blue, %Colour of internal links
  citecolor = blue %Colour of citations
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{12pt}

\begin{document}

\vspace*{1.5\baselineskip}
\begin{flushright}
  \today
\end{flushright}

Dear Editor,

Please find a revised version of
``A spiking neural integrator model of the adaptive
control of action by the medial prefrontal cortex''
attached for your consideration for publication in the
Journal of Neuroscience.

We very much appreciate the reviews
of the original submission.
Considerable effort has been made to address
all issues raised in the reviews.
In order to address these issues,
we have done additional simulations
and revised a significant portion of the text.
Three new figures and several pages
of text have been added
to clarify issues that were unclear
in the original submission,
and to present the results
of the additional simulations.

One notable change in the manuscript,
reflected in the updated title,
is our use of the term ``medial prefrontal cortex,''
or mPFC, instead of ``anterior cingulate cortex,''
or ACC, to characterize the brain area
that was the focus of our study.
This change was made in response to
Reviewer 1, who raised concerns
about functional homologies across species.
Anatomists working on the ACC
consider the prelimbic cortex
(area 32) as a cingulate region
(e.g., studies from the Barbas lab;
\citenoparens{Medalla2009}).
However, there is an unfortunate controversy
about the rat prefrontal cortex,
due in part to the Uylings review
that the reviewer mentioned.
We would like to avoid debate
about this issue in our paper,
and chose to use a more generic term
that is commonly used in rodent studies
to describe the cortical region
in the revised manuscript.

We hope that you find the manuscript
is now suitable for publication
and look forward to hearing from you soon.

Sincerely, \vspace{1em}

Trevor Bekolay, Mark Laubach, Chris Eliasmith

\clearpage
\setlength{\parindent}{1cm}
\setlength{\parskip}{3pt}

\section*{Reviewer 1}

We would like to thank the reviewer for his or her
careful reading of our manuscript
and thoughtful feedback on conceptual
and technical issues in the study.
We have addressed all points raised
in the review,
and we believe that the paper
is very much improved thanks
to Reviewer 1's comments.

\begin{quoting}
  The computational model, named ``double integrator model'' is a
  dynamical system composed of two interacting populations of spiking
  neurons. It is programmed in the Nengo simulation environment, as
  part of the Neural Engineering Framework (NEF; Eliasmith and
  Anderson, 2003) which has the potential to constitute a useful tool
  for neuroscientists to try to simulate, reproduce and better
  understand activities of large cortical networks.

  The description of the model should be strongly clarified and
  extended to enable the reader to precisely understand what is being
  simulated and to distinguish what is part of the general Nengo
  framework and what is specific to the proposed computational model.
\end{quoting}

We appreciate the comments on Nengo and the NEF.
The ``double integrator model'' is originally
from \citetext{Singh2006}.
We have adapted it for the present study
by giving it more inputs, with the most
important being the event that terminates
integration (reward feedback).
Additionally, we have made
a feedforward control network,
and used the double integrator
to control its behavior adaptively.
These distinctions have been made more clear
due to reorganization of the methods section,
and by simplifying several sections of the model.
For example, the \textit{Holding} and \textit{Trigger}
populations only existed in the neural implementation.
We have removed those and ensured that
all neural models are a direct implementation
of their corresponding dynamical systems.
All of the equations governing
the dynamical systems are included
in the revised manuscript.

\begin{quoting}
  More analyses of the results should be provided (dependence to
  parameters, robustness) to better highlight under which conditions
  the model can or cannot reproduce the results - instead of simply
  showing a version of the model that works well.
\end{quoting}

We have now included several numerical simulations
of the dynamical system (Figures 4 and 5)
to clarify the goals of the model,
and under which conditions it meets these goals.
These simulations show how the system changes
as a result of changes to the three free parameters
in the double integrator network equations.
The many parameters affecting the neural implementation
are selected partly at random;
the model performs well and
matches the experimental data
for many different sets of randomized parameters.
We have explicitly listed the distributions
from which we choose neural parameters
in order to make the model easier to reproduce.

\begin{quoting}
  The relation to previous computational models and experimental data
  of the primate ACC and rodent mPFC should be better framed and
  discussed, since it is not clear how the present simulations of one
  particular rat mPFC electrophysiological study can
  relate/complete/replace previous theories of the role of primate ACC
  in cognitive control, as the authors claim in the introduction; the
  discussion should be extended to thoroughly address each raised
  issue.
\end{quoting}

Thanks for this comment.
We must admit that some of the text
in the original manuscript was too bold
with regard to the impact of our model
of delayed response performance
by the rat mPFC.
The text was revised with this in mind,
and we hope that the changes
address Reviewer 1's concerns.

There are many models of ACC/mPFC function
available in the literature.
Given the scope of our study, it was not possible
to address all potential models.
We chose to focus on the recent PRO model
by \citetext{Alexander2011} and included
comparisons of our results
to that model in Section 5.1 and Figure 6.
We have also included Section 5.2 in the discussion
to relate our results to those from
reinforcement learning based approaches like
\citetext{Behrens2007,Alexander2011,Khamassi2012}.

\begin{quoting}
  In the introduction, the objective of the study and its potential
  contribution to important neuroscience debates should be better
  formulated. First, the authors claim that "there are currently no
  mathematical accounts of ACC function that would explain the myriad
  of behaviors modulated by the ACC." But the model presented here
  does not attempt to explain a myriad of behaviors. It only
  reproduces data from a single study (Narayanan and Laubach, 2009).
\end{quoting}

The reviewer is correct, and we should not
have written this text so strongly.
The introduction has been thoroughly
revised in the new manuscript.

\begin{quoting}
  Although I agree that there currently is not any complete model of
  ACC functions, there are several recent computational models that
  have importantly extended the initial models and that the authors do
  not seem aware of (Behrens et al., 2007; Alexander and Brown, 2011;
  Khamassi et al., 2013). I think the manuscript would greatly benefit
  from clearly stating what is new compared to these previous models.
\end{quoting}

As mentioned above, a detailed comparison
to \citetext{Alexander2011} is included
and the other two papers are now cited and discussed.

\begin{quoting}
  Can these models already explain part of the results of (Nayanan and
  Laubach, 2009)? Even if yes, I think that the dynamical aspect of
  the present model provides something more that should be better
  highlighted. For instance, I think that both the models of Alexander
  and Brown 2011 and Khamassi et al. 2013 could account for the ACC
  activity reported in Narayanan and Laubach 2009 that "change
  significantly after errors, and appear to contain information that
  an error occurred". This should be discussed.
\end{quoting}

Reviewer 1 is correct that the models presented in
\citetext{Alexander2011} and \citetext{Khamassi2012}
would be able to account for some elements
of the \citetext{Narayanan2009} data.
However, we believe that our ability
to generate spike data that can be
directly compared to the experimental data
is a unique contribution,
in additional to the dynamical aspect
that the reviewer notes.
While the information
being tracked is similar between
all of these models
(as shown in the revised manuscript's Figure 6),
we interpret that information
from a dynamical systems perspective
which provides an alternative hypothesis
of how that information might come about
(as Reviewer 1 states).
We have highlighted the dynamical aspect of the model
(particularly in Section 5.2, but also throughout),
and thank Reviewer 1 for pointing out
the similarities between these models.

\begin{quoting}
  It is not true that since ``no existing models have been implemented
  with spiking neurons [...] no models can be analyzed with the same
  methods used to analyze single neuron recordings.'' Simulations of
  simple non-spiking neural network models or even of algorithmic
  models can be analyzed with methods that consider neurons' firing
  rate in different time bins.
\end{quoting}

We respectfully disagree on this point.
While it is true that
our analyses could be performed with
firing rates in time bins,
there are more detailed analyses that can only
be performed with raw spike data.
For neural network models with
real-valued output from each neuron,
a representation scheme must be used
to represent that real-valued output
with spiking neurons;
firing rate alone is rarely sufficient.
For rate models that emit firing rates
and use a representation scheme,
the temporal dynamics in a spiking model
are significantly different;
a neural integrator, for example,
requires about twice as many
spiking neurons than rate neurons
to obtain similar performance.

In short, there are a variety
of differences between rate and spiking models
that require additional analyses
or ``translations'' of rate models
for comparison with spike data,
which are not required of spiking models.

\begin{quoting}
  I think the important thing here is that some characteristics of
  recorded neural activities can be seen as emerging properties of a
  dynamical system which a spiking neural network model enables to
  reproduce.
\end{quoting}

We agree with this statement,
and have attempted to highlight this
in the revised discussion.
As a general statement on neural modeling,
we believe that the aim should not be
to reproduce neural activities,
but instead to produce a working mathematical model.
By implementing the model in a spiking neural system,
we ensure that the model is robust
to the temporal dynamics of spiking neural networks,
and are able to compare directly
to recorded neural activities.
In this sense, the neural activities can be thought of
as ``emerging'' from the neural realization
of the mathematical model.

\begin{quoting}
  I think there is a major issue that needs to be solved concerning
  the acceptability or non-acceptability of considering that the mPFC
  (more precisely the prelimbic cortex (PL)) in rodents is the same
  thing as the ACC in primates. The state of the art of related
  previous computational models provided in the introduction only
  discusses models of ACC function in primates. It does not discuss
  any computational model of PL in rodents (e.g. see Koene and
  Hasselmo, 2005; Martinet et al., 2011). Classically, the rodent PL
  is considered as subserving homologous functions to the dorsolateral
  prefrontal cortex (DLPFC) in primates (Uylings et al., 2003). But
  people agree that the PL in rodents is less differentiated than the
  DLPFC in primates, and PL may integrate functions homologous to both
  the primate DLPFC and ACC. However, it should be taken into
  consideration that the more dorsal part of mPFC in rodents (named
  ACd/Fr2 in the Paxinos) could also partly integrate some functions
  homologous to the ACC in primates. Globally the authors should give
  more justifications. Cite for instance Seamans et al. (2008).
  Nevertheless, it is not straightforward to draw conclusions about
  ACC functions in primates from a model that has been calibrated to
  reproduce electrophysiological activities from a single study in the
  rodent PL (i.e. Narayanan and Laubach, 2009). Maybe the paper would
  benefit from a rewriting of the Introduction mainly describing
  previous experimental results and computational models of PL in
  rodents (although still briefly mentioning the implications for the
  understanding of the role of ACC in primates), and discussing in
  details the possible homologies with the ACC in primates in the
  Discussion.
\end{quoting}

The terms used for describing the medial frontal cortex
across species continues to be an issue for us (ML)
and many other researchers.
The original manuscript used
the term ``anterior cingulate cortex,'' or ACC.
Our rationale for this was that anatomists
working on the primate ACC
have commonly included the prelimbic cortex
(area 32) in the ``ACC''
(for example, studies from the Barbas lab;
\citenoparens{Medalla2009}).
However, there is an unfortunate controversy
about the rat prefrontal cortex,
due in part to the Uylings review
that the reviewer mentioned.
One of us (ML) has spent much time
discussing this issue
in conferences and workshops
(see \citenoparens{Laubach2011}).
We would like to avoid debate about this issue
in our paper, as it has little to do
with the validity of a double integrator network
as a candidate model
for explaining computations
performed by the rodent PFC
in controlling timed actions.
We chose to use a more generic term
(medial prefrontal cortex, or mPFC),
which is commonly used in rodent studies,
to describe the cortical region in the revised manuscript.
We have also limited the scope of our claims.

Following the reviewer's suggestion,
we added text to the Introduction
to review the series of studies
done by the Laubach lab
that led to collaboration
with the Eliasmith lab
on the double integrator network.

\begin{quoting}
  The presentation of previous results obtained by Narayanan and
  Laubach 2009 is too short and is not described with enough details
  to enable the reader to fully understand without reading the
  original paper. The authors should dedicate more space to describe
  the results displayed in Figure 2 and to clearly explain the
  interpretation they make. In particular, what does figure 2F tell
  us? Also, we would like to see on figure 2 the onset of the correct
  and error related PCS at the time of the feedback.
\end{quoting}

Thanks for raising this issue.
We agree that more should have been done to make Figure 2 clear,
and panel F was not sufficiently explained
in the original paper.
A paragraph was added to Section 2.5.1
to discuss the plots in panel F,
and the timing of reward feedback
has been noted in the figure.
Importantly, we note that
the analysis was only done on correct trials,
and so feedback was at the same time
for bot the post-correct and post-error trials.

\begin{quoting}
  The model should be described in more details both for specialists
  who would like to reimplement it themselves and for experimentalists
  who will read the paper. First, the terms used in equation (6)
  should be better explained. Second, it is not always clear which
  computations are specific to the proposed computational model and
  which computations are general to the NEF framework to implement any
  dynamical system and ``any neuron model''. To make the message
  clearer, I think it is important here to stick to the description of
  the proposed computational model, and to mention in a clearly
  separated paragraph what describes properties of the NEF (and of
  Nengo) that could be interesting to a reader that would like to
  program his own model.
\end{quoting}

Some confusing and unnecessary aspects of the model
have been simplified.
For example, the \textit{Holding} and \textit{Trigger} populations
were removed.
They originally existed in order to extend
the model to more general situations
in which motor actions may be
contingent on body state
through visual or proprioceptive feedback;
however, we have instead constrained the model
and linked the results only to the simple RT task
in order to simplify explanation
and avoid overgeneralizing the model
to tasks we have not modeled.

We have made several improvements to
the methods sections on
the dynamical systems and the NEF.
Included in those improvements
is a rewritten Equation (6)
that now uses matrix notation
to make the mathematical operation
(least-squares minimization) more clear.
While it was also the case
in the original paper
that NEF-specific equations
were only listed in one section,
we hope that with these improvements
it is more clear that all of the equations
in Section 2.4 are specific to the NEF,
and all equations in other sections
are part of the dynamical systems
that are simulated both directly
and implemented in spiking networks with the NEF.

Due to the lack of supplementary material,
it is difficult to include
a sufficiently detailed explanation of the NEF
to make it easily reproducible
from this paper alone.
We have opted for a comprehensive
explanation that includes
all of the equations necessary
to reproduce the results,
but with shorter explanations
than can be found
in other resources on the NEF.
However, we believe that the open source nature
of Nengo and the scripts that we use
to generate the model
make it possible for specialists
to reproduce the model.

\begin{quoting}
  The authors should avoid using the term "model" for many different
  things, since this can be very confusing for experimentalist readers
  (i.e. the computational model; "any neuron model"; "for each
  instance of each model"; ...).
\end{quoting}

The many levels of modeling that are used in this study
makes an overuse of the word ``model''
difficult to avoid;
we thank the reviewer for pointing this out.
We have attempted to minimize
the use of the word model.
Specifically, we refer to the dynamical systems
(the double integrator, cue-responding,
and adaptive control systems) and their neural
implementations as ``networks.''
All uses of the term ``model'' are prefaced
with a identifying term such as
``neuron'' or ``network'' in order
to avoid ``model'' losing its meaning
as a mathematical idealization.

\begin{quoting}
  The authors mention a mechanism for switching between an aggressive
  timing strategy and a conservative cue-response strategy. What is
  the programmed criterion for switching? Please write the used
  equations.
\end{quoting}

Thanks to Reviewer 1 for pointing out this glaring omission.
We have included the switching criterion as Equation (5)
in the revised manuscript.

\begin{quoting}
  It is important to provide analyses of the strengths and weaknesses
  of the proposed computational model, so that the reader can really
  understand what enables and what does not enable the model to
  succeed in reproducing the experimental data. What is the
  sensibility to changes in parameter values? Could the authors
  illustrate the quality of reproduction of experimental data as a
  function of different parameter sets?
\end{quoting}

We have attempted to explain the meaning of the parameters
through the newly added text in Section 3.1,
and its associated analyses in Figure 4.
We have also given more context to these parameters
in the explanations of the dynamical systems
in Sections 2.2 and 2.3.

Aside from the three parameters we set,
the model's success is dependent solely
on the ability of the integrative populations
to retain their represented values over time.
In our case, model fits are generally good
(Pearson $R^2 > 0.7$)
in any case when the random neural parameters
produce good integrators.
The ALIF neuron introduces additional dynamics
that make creating neural integrators
more difficult than
neural integrators made up of LIF neurons,
but these difficulties can be overcome
with larger populations.
The original double integrator paper
\cite{Singh2006}
discusses the robustness
of NEF-created neural integrators
in the double integrator context,
and \citetext{Eliasmith2005} discusses this issue
for many NEF-constructed attractor networks.
More detailed analyses can be found
in \citetext{Eliasmith2003}.

\begin{quoting}
  Related to this issue, on page 11, an important question is how much
  of parameter tuning was necessary to obtain the same PCs as
  experimental data? It is good to reproduce the experimental results.
  But it is even better to show analyses of different regimes or
  states the model can produce with different parameters, and to
  highlight which of these states or regimes may the experimental data
  correspond to.
\end{quoting}

We now discuss in the paper
how each parameters affects the model,
and show simulations with different sets of parameters
(Table 1 describes the ALIF parameters;
Table 2 and Figure 4 describe the
dynamical system parameters).
Section 3.1 discusses the results
of the simulations
and lists the parameters values
that were chosen.
Importantly, the parameter choices
did not require fitting,
and for most parameters,
a range of values would produce
functionally identical results.

For the experimental model fits in Figure 7,
several models were generated and the best
experimental fit was chosen.
It is possible to show the
average fits for
several instantiations of the model,
but this is not very informative because
increasing the number of neurons in the model
produces networks that fit the data well
nearly all the time,
despite many randomly generated parameters.
As noted above, the performance of the integrators
is the critical factor to the performance
of the entire model.

\begin{quoting}
  In relation to the previous point, it seems quite ad hoc and looks
  like a strong simplification to hand tune the model so that "reward
  delivery [...] resets the system to a starting state" while "an
  error trial [drives it] to a low state". Shouldn't different
  reactions of the model to correct versus incorrect trials emerge
  through learning (such as in reinforcement learning-based models of
  ACC; e.g. Alexander and Brown, 2011; Khamassi et al., 2013)? Or
  shouldn't they at least produce symmetrical effects on the model
  that it should exploit to produce different reactions? Isn't this a
  strong help provided a priori to the model to make it react in the
  expected way? Maybe I am wrong to think that this is a
  simplification. Maybe this is equivalent to the priors put in the
  cited previous computational models. But this should be at least
  discussed by the authors. For instance, doesn't such simplification
  prevent generalizations of the model to tasks involving a sequence
  of rewards and/or punishments (e.g. Tanaka et al., 2004)?
\end{quoting}

This is an important point that identifies
the central difference between
models created with the NEF
and many other computational modeling approaches,
including reinforcement learning-based approaches.
What Reviewer 1 calls hand tuning and simplification
is essentially our dynamical systems hypothesis
of how to solve the reaction time task
with a control system operating in continuous time.
While it may seem a simplification
for reward to drive the system to the origin
and error to drive it to a low part of the state space,
these reflect the operations
that are possible (and, indeed, natural) in dynamical systems.
We use these operations to implement a dynamical system
that performs the task
as we hypothesize the rodent performs the task.
We can then implement that dynamical system
in a network of spiking neural populations
through the principles of the NEF.
This spiking neural network models
the neural circuit of a rodent
that has already learned the task;
going from random connectivity to this network
is a separate branch of research.
We believe that providing the end-point of learning
is a great help to modelers
aiming to learn this kind of neural circuit
(and have shown an example of this
in other work; e.g., \citenoparens{MacNeil2011}).

In other words, the NEF provides three biologically plausible
principles for network construction
through which we can produce a neural circuit.
We have presented a way in which those principles
can be used to generate a network
that performs the simple RT task.
We have attempted to
emphasize these consideration
in various parts of the revised discussion,
most notably Sections 5.1 and 5.2.

\begin{quoting}
  Some simulation results are too briefly presented -- sometimes
  making it difficult to understand for the reader -- and some are not
  analyzed with enough deepness. On page 10, the results displayed in
  figure 4 should be better explained in the main text. A few lines
  below, in contrast to what the authors wrote, it is not clear for
  the reader from these trajectories that the time of the cue can be
  predicted. During correct trials, the cue has already occurred to
  make ACC in the right starting position for the next trial. Please
  explain what you mean here.
\end{quoting}

Simulations of the dynamical system are expanded
and simplified, as discussed below.
We have included two new figures (4 and 5)
that aim to help with understanding
how the simulations work under different sets of parameters.
More detailed equations throughout Sections 2.2 and 2.3
are also included in the revised manuscript.

\begin{quoting}
  The authors should show, either on figs 4 and 5 or on supplementary
  figures, correct, premature and late trajectories starting from the
  lower-left starting position (following an error).
\end{quoting}

Figure 4 in the revised manuscript now presents trajectories
from many different initial conditions.
It also shows how the result of the last trial
determines the initial conditions
of the next trial.

\begin{quoting}
  Can the states be decoded in the same manner when starting from such
  a different position?
\end{quoting}

The state can be decoded,
but the cue cannot be accurately predicted
unless the system starts at the origin
under normal system behavior.
This is why cue prediction does not occur
in post-error trials;
the system starts in the lower part
of the state space instead of the origin.
This is shown and expanded on in Figure 4
and Section 3.1.

\begin{quoting}
  On page 11, the authors show that the double integrator model has
  the same PCs as the experimental data. Could the authors also show
  whether the cue-responsive model also produces similar PCs as
  experimental data?
\end{quoting}

The cue-responsive network does not contain any integrative
populations, so it is highly unlikely that
the PCs would be the same as the experimental data.
However, there is an open question
(which was raised by Reviewer 2)
as to whether other network structures
would result in the same principal
components as the double integrator network.
In order to address this question,
we have added Figure 8 in the revised manuscript,
which shows that
a single integrator does not produce
similar PCs as experimental data.

\begin{quoting}
  In Narayanan and Laubach (2009), premature and late cases are
  combined in a single PC1 and PC2 couple of curves for "post-error".
  Do the PCs of the original data (from Narayanan and Laubach 2009)
  still look like the model's curves when premature and late cases are
  separated?
\end{quoting}

In \citetext{Narayanan2009}, the PC analysis
was only done using correct trials.
Error trials were not included in the analysis.
PCA done on premature trials
deviate from those based on correct trials
due to the rats responding early in the foreperiod.
PCA on late trials do not differ much
from those on correct trials,
but lack modulation during the normal period
when rats release the lever
at the end of the foreperiod.

\citetext{Narayanan2008},
which was mistakenly not cited in the previous
version of the paper, reported effects
for premature and late responses,
and showed that single units
(not population activity)
tracked outcomes during
the intertrial interval.

Importantly, while neural activity
tracked outcomes during the intertrial interval,
that activity did not differentiate
between the type of error that occurred
\cite{Narayanan2008}.
Therefore, we combine post-premature
and post-late trials when analyzing
the experimental data
in order to increase
the number of trials being analyzed,
and therefore increasing
the statistical power of the analysis
For the simulated data,
we analyze the two separately
in order to show that the model also
does not differentiate between
the two cases;
since we can simulate as many
trials as we want,
we do not need to combine
the types of error trials
in order to increase statistical power.

\begin{quoting}
  Could the authors explain how late trials can occur in the adaptive
  control model nearly as often as the other model while the former
  anticipates and responds faster than the latter?
\end{quoting}

Late trials can occur in exactly the same way
as in the cue-responding model post-error,
as the adaptive control model takes
a cue-responding strategy in that case.
Like in the cue-responding case,
late trials often correspond to non-responses,
in which the release should have occurred,
but release sensitive neurons were
not active for long enough to actually
effect the full lever press.
These errors occur at approximately
the same frequency in the cue-responding
and adaptive control models.

\begin{quoting}
  Can the model also reproduce activities during the second version of
  the task in Narayanan and Laubach 2009 where cues are presented only
  on half of the trials?
\end{quoting}

Because the cue is predicted by the double integrator network,
the model can accomplish the task even if the cue
is not actually present.
The model would not change
or produce different PCs in this case,
so we did not discuss the second version of the task
in order to keep the focus of the paper clear.

\begin{quoting}
  Do the simulation results really demonstrate that the ACC cannot
  inhibit the motor command? (Page 9). If ACC is x1+x2, then what
  brain structure correspond to H+T+P/R units? If it is also ACC, the
  authors need to first validate their ability to reproduce recorded
  ACC activity. If not, then how does the model prove that ACC is not
  inhibiting motor responses? A convincing demonstration would have
  been to simulate the two alternative hypotheses and show that one
  fails to reproduce some experimental evidence. Otherwise, the
  authors should not sell this part of the work (page 9) as a strong
  demonstration, but rather just say that with this ACC model, it is
  not necessary to exert inhibition on motor cells to work.
\end{quoting}

We have removed sections
discussing the motor inhibition hypothesis
from the paper.
We thank Reviewer 1 for pointing out
the weaknesses of this argument;
we believe that the revised discussion
proposes more concrete predictions
that come from the model.

\begin{quoting}
  The prediction extracted from the model on page 12 ("our model
  predicts that fewer premature releases would occur") appears like a
  trivial prediction. It seems to the reader that there is no need of
  a computational model to predict this. Either expand on this
  prediction or give arguments to convince the reader that this is not
  a trivial prediction.
\end{quoting}

We agree that this prediction is trivial,
and have removed it from the revised manuscript.

\begin{quoting}
  The model should not only be evaluated in its ability to reproduce
  neurophysiological recordings, but also in its ability to reproduce
  rat behavior simultaneously recorded. For instance, on page 10, it
  is stated that "if the previous trial was an error, the double
  integrator will not be following the trajectory that predicts the
  time of the cue, and therefore will not cause a lever release." Is
  it what rats do in Narayanan and Laubach (2009)? Please quantify
  such behavior and other behaviors of the model and compare with rat
  behavioral data.
\end{quoting}

Unfortunately, it is impossible to determine if rodents
are predicting the time of the cue given their behavior alone;
we could estimate this by analyzing
the reaction times on correct trials
following an error,
but due to amount of variability in reaction times,
there are not enough instances of post-error correct trials
to definitively state that reaction times are slower
on these trials compared to post-correct correct trials.

We provide a summary comparing
the behavioral statistics of
the spiking cue-responding and adaptive control networks
to the experimental subjects in Figure 10.
We believe this is the most appropriate evaluation
given the data available.

\begin{quoting}
  The manuscript would be strengthened if the authors could provide a
  series of testable experimental predictions that would allow to
  further test/refute the proposed computational model. In particular,
  what further experiments are needed to decide between the two
  mentioned alternative hypotheses about ACC inhibiting motor
  responses or this inhibition being elsewhere?
\end{quoting}

We have added Section 5.3 listing several predictions.
Since we do not discuss the motor inhibition hypothesis,
we do not propose an experiment relating to this hypothesis.
However, we do make predictions about
how neurons might respond in tasks
with different temporal dynamics,
and discuss experiments that can be done
to confirm or refute our predictions.

\begin{quoting}
  The discussion about the necessity to differentiate between
  adaptability due to recurrent activity and adaptability due to
  synaptic weight changes is very interesting and should be further
  elaborated and extended. In particular, previous ACC models rely on
  reinforcement learning mechanisms that would induce adaptability due
  to synaptic weight changes (Alexander and Brown, 2011; Khamassi et
  al., 2013). Here the model produces adaptability due to recurrent
  activity. It would be useful to discuss whether some of the results
  produced by these previous models may be also explained by recurrent
  activity. If yes, could the authors think of an experimental
  protocol that would enable to decide which of the alternative models
  is right?
\end{quoting}

We agree that differentiating
between adaptability
through recurrent activity
and through synaptic weight changes
is an important issue
that we think will become increasingly important
as more complex mPFC models are proposed.
We have added Section 5.2 to discuss
this differentiation.

In the revised discussion, we have proposed two
new experiments that could allow for
direct comparisons between neural data
and the simulated networks.
First, one could record as rats are initially trained
to perform the task and compare
the dynamics of the mPFC (and motor cortex)
with a double integrator network.
A challenge for the theory side would be
to evaluate various types of RL algorithms
that might enable a match
between theory and data.
Second, one could train rats to perform
a variant of the simple RT task
using multiple foreperiods,
and test the rats in sessions with
interleaved and blocked foreperiods.
If the mPFC neurons
and double integrator neurons track
the expected timing of action,
as we predict, then there would be
shifts in the temporal expectations
following each switch in block
(e.g., from 0.5 to 1.5 second foreperiods
over blocks of 30 trials).
To our knowledge,
no group has carried out such studies.
However, they could allow for
direct comparisons between
brain activity and theory.

\begin{quoting}
  The discussion about possible extension of the model to account for
  data in the Stroop task should be discussed in relation to previous
  models. What would such extension of the authors' model add to
  previous simulations with the model of Dehaene and Changeux (1998)?
  Globally, the discussion should be extended to thoroughly address
  each raised issue.
\end{quoting}

We have removed the possible extension of the model
to the Stroop task and instead highlighted
the model's specificity to the simple RT task.
However, we have also highlighted in Section 5.2
that the process of creating
dynamical system for performing the simple RT task
could also be done for many other tasks
and implemented using the same methods.

\section*{Reviewer 2}

We would like to thank Reviewer 2
for noting the interesting point that
the double integrator model and the published
mPFC recordings share the same
population statistics, and for inquiring
about some details in the neural circuit model.
We believe that we have been able to
address all of his or her concerns,
and, as above for Reviewer 1,
the manuscript is much improved
thanks to his or her suggestions.
In what follows,
we detail our responses to each comment.

\begin{quoting}
  I think the most compelling result in the paper is in figure 6 --
  i.e. that the network has the same principal components as the
  neural data. One question I have is to what extent this result is
  unique to a double integrator or could other networks produce these
  kind of PCs? My feeling is you have to have something like a double
  integrator but I would like to see some control simulations that
  show that. So, for example, would a single integrator network have
  those PCs? Would a random network with the same inputs have those
  PCs?
\end{quoting}

This is an interesting point,
especially when using principal component analysis
on filtered and Z-scored spike trains.
To address this issue, we added
principal component analysis results
from a single integrator network (Figure 8),
and show that it does not have
the same PCs as the double integrator
or the neural recordings.

\begin{quoting}
  I'd like to see slightly more in section 2.3.2. In particular what
  are the exact update equations for the network? In particular what
  is the exact form of $G$ what are the values of $J_{bias}$ etc ... I
  realize these details are in the referenced paper and book but I
  think it is appropriate to have them in an abridged form here.
\end{quoting}

The update equation for $G$ has been included in Section 2.4
(which was formerly Section 2.3.2),
and a table of the parameters used
(that discusses how $J_{bias}$ is determined)
has been included as Table 1.

\begin{quoting}
  I might have missed this - but how many neurons are in the model?
  You say you sample 174 but I'm not sure how many are in the actual
  model?
\end{quoting}

Thanks for catching this issue.
The simulated networks had 1200 neurons;
this has been stated in the revised text
and in Figure 3.

\begin{quoting}
  Regarding the fact that the behavior of the model is slightly too
  good (not enough errors) -- you mention adding noise to the model
  might solve this problem. I think it would be worth reporting
  whether it does in practice. Neural integrators can be fickle beasts
  so I don't think this is a given.
\end{quoting}

Our original mention of noise was in reference
to the spike density functions plotted
in (now) Figures 7 and 8.
While the principal components matched closely,
the spike densities revealed that
the simulated data is much cleaner
than the experimentally recorded data.
As Reviewer 2 suggests, we added noise
($J_\eta$, now included in equations
in Section 2.4)
and recreated Figure 7 (formerly Figure 6).
As can be seen in the revised manuscript,
the simulated spike densities now very closely resemble
the experimental spike densities,
while still maintaining a very close
principal component fit.

In regards to the model performing too well behaviorally,
this is not an issue of the noisiness of the system,
but on other modeling decisions.
In particular, we could have varied $\beta$
in order to obtain more premature errors,
which would match the behavior
of experimental subjects more closely.
However, rerunning the simulations
that are summarized in Figure 10
takes several days.
Since having a perfect fit
to the experimental performance
is not necessary to match other quantities
like reaction times,
we chose an acceptable value for $\beta$
based on direct simulations
of the dynamical system and
did not vary it while simulating the spiking implementation.

\begin{quoting}
  I think it would be nice if you could generate some testable
  predictions of the model. For example, can you say anything about
  the character of neurons that contribute to each of the principal
  components ... is there any difference between them in the model in
  terms of connectivity (are PC2 neurons more highly connected than
  PC1 neurons) and sign (are PC1 neurons more likely to be
  inhibitory). What about microstimulation or inactivation? Similarly
  for the holding, trigger and press release neurons - how would you
  know if you were recording them? As I understand it they might not
  look like figure 8 because that's the decoded signal not firing
  rate. These are just suggestions and I will leave it to the authors
  to try and find something - but the nice thing about a model is you
  can interrogate it and I think it would improve the paper to put
  down some firm predictions that can be tested.
\end{quoting}

Thanks for this suggestion,
which was also raised by Reviewer 1.
Testable predictions have been added in Section 5.3.

\begin{quoting}
  Will the Nengo scripts be available through J Neuroscience? I'm not
  quite sure what the policy is for this kind of material but it would
  be nice if it were available with the article as the github may
  change over time ...
\end{quoting}

A link to the files used
for the submitted version of the manuscript
is provided in the revised manuscript.
This link will always point to the
version of those files
used to generate the revised manuscript
even if changes are subsequently
made to those files.
Additionally, we have also included
at that link a complete copy
of the Nengo simulator
used to produce the model,
in case changes to Nengo also
affect model behavior.
With this, any computing environment
with Java should be able to
run the simulations,
and a computing environment
with Python and several open source libraries
(listed with the analysis scripts)
should be able to run
the analysis and generate
the plots shown in the paper.

\bibliographystyle{namedplus}
\bibliography{jneurosci2013}

\end{document}